---
title: "clean Haute Borne Windfarm Data"
author: "Max Davis"
date: "6/29/2019"
output:
  html_document: default
  pdf_document: default
---

Load packages and libraries

```{r}

install.packages("readr")
install.packages("data.table")
```

```{r}
library("readr")
library(dplyr)
library(tidyr)
library(ggplot2)
library(data.table)
```

Read data for full Haute Born Wind farm from 2017 to 2020, and the data description file or metadata using the fread() function with data.table = FALSE:

```{r}
haute_all_2017_2020 <- fread("~/Desktop/DATA/Haute Borne Wind Farm Data/la-haute-borne-data-2017-2020.csv", data.table = FALSE)

data_description <- fread("~/Desktop/DATA/Haute Borne Wind Farm Data/data_description.csv")
```


make tibble:

```{r}

haute <- tbl_df(haute_all_2017_2020)

```

run functions that change problematic variable names (those that happen containing gsub two-letter name) before changing the rest.

```{r}

colClean27 <- function(x) { colnames(x) <- gsub(pattern = "P", "Active_power", colnames(x)) ; x}
colClean20 <- function(x) { colnames(x) <- gsub(pattern = "Va", "Vane_position", colnames(x)) ; x}


  haute_clean_1 <-
  haute %>%
  colClean27() %>%
   colClean20()

```

Create functions for each variable name that subs in long descriptions

```{r}

colClean <- function(x) { colnames(x) <- gsub(pattern = "Q", "Reactive_power", colnames(x)) ; x} 
colClean2 <- function(x) {colnames(x) <- gsub(pattern = "Va2", "Vane_position_2", colnames(x)) ;x} 
colClean3 <- function(x) { colnames(x) <- gsub(pattern = "Ws2", "Wind_speed_2", colnames(x)) ; x}
colClean4 <- function(x) { colnames(x) <- gsub(pattern = "Ws", "Wind_speed", colnames(x)) ; x}
colClean5 <- function(x) { colnames(x) <- gsub(pattern = "Dst", "Generator_stator_temperature", colnames(x)) ; x}
colClean6 <- function(x) { colnames(x) <- gsub(pattern = "Git", "Gearbox_inlet_temperature", colnames(x)) ; x}
colClean7 <- function(x) { colnames(x) <- gsub(pattern = "Ot", "Outdoor_temperature", colnames(x)) ; x}
colClean8 <- function(x) { colnames(x) <- gsub(pattern = "Yt", "Nacelle_temperature", colnames(x)) ; x}
colClean9 <- function(x) { colnames(x) <- gsub(pattern = "Nf", "Grid_frequency", colnames(x)) ; x}
colClean10 <- function(x) { colnames(x) <- gsub(pattern = "Nu", "Grid_voltage", colnames(x)) ; x}
colClean11 <- function(x) { colnames(x) <- gsub(pattern = "Rm", "Torque", colnames(x)) ; x}
colClean12 <- function(x) { colnames(x) <- gsub(pattern = "Wa_c", "Absolute_wind_direction_corrected", colnames(x)) ; x}
colClean13 <- function(x) { colnames(x) <- gsub(pattern = "Na_c", "Nacelle_angle_corrected", colnames(x)) ; x}
colClean14 <- function(x) { colnames(x) <- gsub(pattern = "Gb2t", "Gearbox_bearing_2_temperature", colnames(x)) ; x}
colClean15 <- function(x) { colnames(x) <- gsub(pattern = "Gost", "Gearbox_oil_sump_temperature", colnames(x)) ; x}
colClean16 <- function(x) { colnames(x) <- gsub(pattern = "DCs", "Generator_converter_speed", colnames(x)) ; x}
colClean17 <- function(x) { colnames(x) <- gsub(pattern = "Wa", "Absolute_wind_direction", colnames(x)) ; x}
colClean18 <- function(x) { colnames(x) <- gsub(pattern = "Ds", "Generator_speed", colnames(x)) ; x}
colClean19 <- function(x) { colnames(x) <- gsub(pattern = "Cm", "Converter_torque", colnames(x)) ; x}
colClean21 <- function(x) { colnames(x) <- gsub(pattern = "Ya", "Nacelle_angle", colnames(x)) ; x}
colClean22 <- function(x) { colnames(x) <- gsub(pattern = "Rs", "Rotor_speed", colnames(x)) ; x}
colClean23 <- function(x) { colnames(x) <- gsub(pattern = "S", "Apparent_power", colnames(x)) ; x}
colClean24 <- function(x) { colnames(x) <- gsub(pattern = "Cosphi", "Power_factor", colnames(x)) ; x}
colClean25 <- function(x) { colnames(x) <- gsub(pattern = "Rt", "Hub_temperature", colnames(x)) ; x}
colClean26 <- function(x) { colnames(x) <- gsub(pattern = "Ba", "Pitch_angle", colnames(x)) ; x}

colClean28 <- function(x) { colnames(x) <- gsub(pattern = "Va1", "Vane_position_1", colnames(x)) ; x}
colClean29 <- function(x) { colnames(x) <- gsub(pattern = "Ws1", "Wind_speed_1", colnames(x)) ; x}
colClean30 <- function(x) { colnames(x) <- gsub(pattern = "Db2t", "Generator_bearing_2_temperature", colnames(x)) ; x}
colClean31 <- function(x) { colnames(x) <- gsub(pattern = "Db1t", "Generator_bearing_1_temperature", colnames(x)) ; x}
colClean32 <- function(x) { colnames(x) <- gsub(pattern = "Gb1t", "Gearbox_bearing_1_temperature", colnames(x)) ; x}
colClean33 <- function(x) { colnames(x) <- gsub(pattern = "Rbt", "Rotor_bearing_temperature", colnames(x)) ; x}
colClean34 <- function(x) { colnames(x) <- gsub(pattern = "Pas", "Pitch_angle_setpoint", colnames(x)) ; x}
```  

Run each function from previous code chunk on haute_clean_1 and store as haute_clean:

```{r}
haute_clean <-

  haute_clean_1 %>% 
  colClean() %>%
  colClean2() %>%
  colClean3() %>%
  colClean4() %>%
colClean5() %>%
colClean6() %>%
colClean7() %>%
colClean8() %>%
colClean9() %>%
colClean10() %>%
colClean11() %>%
colClean12() %>%
colClean13() %>%
colClean14() %>%
colClean15() %>%
colClean16() %>%
colClean17() %>%
colClean18() %>%
colClean19() %>%
colClean21() %>%
colClean22() %>%
colClean23() %>%
colClean24() %>%
colClean25() %>%
colClean26() %>%
colClean28() %>%
colClean29() %>%
colClean30() %>%
colClean31() %>%
colClean32() %>%
colClean33() %>%
colClean34()
```

```{r}
View(haute_clean)
```
Make a .csv file in the working directory:

```{r}
write.csv(haute_clean, file = "haute_clean.csv")
```

