---
title: "clean Haute Borne Windfarm"
author: "Max Davis"
date: "6/29/2019"
output: html_document
---

Load 3 datasets:

la-haute-borne-data-2017-2020.csv
data_description.csv
limited_la-haute-borne-data-2017-2020.csv (import without converting strings to factors to get names as characters)

```{r}

library(dplyr)
library(tidyr)
library(ggplot2)
```

make tibble:

```{r}
haute <- tbl_df(limited_la.haute.borne.data.2017.2020)
hautenamesource <- tbl_df(char_limited_la.haute.borne.data.2017.2020)
```

```{r}
hautenames <- as.character(hautenamesource[1, 1:138])
```


Make first row into column names:
```{r}
colnames(haute) <- hautenames
View(haute)

```
 remove first row of column names:
```{r}
haute <- haute[-1,]
View(haute)
```

FUTURE: Make column names readable by changing abbreviation to longer name.

old.name.part <- as.character(data_description[, "Variable_name"])
part.replacement <- as.character(data_description[, "Variable_long_name"])
str(la.haute.borne.data.2017.2020[1,])


View histogram of power output





ggplot(haute, aes(x = Ds_avg, y = P_avg))








